\boldmath Primordial Black Hole Formation in a Dust Bouncing Model

### 2024-11-19 referee_response.tex

              \item Indeed, if the sound speed were to approach zero, we would encounter
                    a divergence problem. This is why we limit ourselves to small but
                    non-zero equations of state for dust, which is in agreement with the
                    literature. Regarding the $\zeta$ field, there are always field
                    configurations that would lead to non-perturbative behavior.
                    However, in the present context, given the initial conditions in
                    equation 3.32, we are considering the adiabatic vacuum for which
                    $\zeta \propto 1/\sqrt{z^2c_s} \propto \sqrt{c_s}$. Therefore, terms
                    of the form $z^2c_s^2\zeta^2$ are $\propto c_s$ while $\zeta^3
                        \propto \sqrt{c_s}^3$ This means that the perturbative expansion is valid
                    for small sound speeds, as discussed in
                    Ref.~\cite{vitenti2012large}.

                    We have reorganized Sections 4.3 and 4.4 and considered a very
                    small, but non-zero, equation of state for dark matter to analyze
                    the pressureless case.
          \end{enumerate}

    \item

          \begin{enumerate}
              \item We improved Sec.~2.2 by adding a more detailed explanation of the
                    Schutz formalism and the fluid Lagrangian. We also included a
                    discussion of the lapse function and how it can be defined as a
                    posteriori for simplicity.

              \item In the improved Sec.~2.2, we have added a brief discussion of the
                    meaning of the effective energy density and pressure in the context
                    of the quantum corrections. In Ref.~\cite{fluidgeral}, the authors
                    show that the energy density and pressure appearing in the perturbed
                    equations are only the classical ones computed with the Bohmian
                    trajectory. Therefore, the equation-of-state parameter is always
                    $w=p_d/\rho_d$.

              \item In the revised Sec.~2.2, we have included the effective equation
                    corresponding to the second Friedmann equation, now presented as Eq. 2.17:
                    $-2\dot{\bar{H}} = \bar{\rho}(1+w) - 2\rho_q.$
                    Notably, the right-hand side of this equation approaches zero around the
                    bounce. However, as discussed previously in Ref.~\cite{fluidgeral} it
                    has been shown that the energy density and pressure in the second-order
                    Lagrangian for the perturbations are only the classical components
                    evaluated along the Bohmian trajectory. Consequently, this term is never
                    zero, ensuring that the model is free from ghost instability issues.
              \item In Ref.~33 the author chooses a particular form for the wave
                    function at the bounce. This is indeed not an initial condition, but
                    a choice made to simplify the analysis. The same wave-function
                    propagated to $T\to-\infty$ would have an extra phase factor
                    proportional to the $q^2$, exactly like a wave function of a
                    free Gaussian package in 1d. One can always start with a different
                    wave function and define the bounce as the time where the phase
                    factor proportional to $q^2$ is zero. Again, this is equivalent to a
                    free Gaussian package in 1d. Thus, in short, it has to start at a large
                    but finite negative time, with a Gaussian wave-function containing a
                    phase factor proportional to $q^2$. Then, the bounce is defined as
                    the time when this phase factor is zero. We have included a brief
                    discussion of this point in the revised manuscript.
              \item In our approach, we use a perturbative method: the background is
                    solved first, and the Bohmian trajectory is derived from the
                    background wave function. Perturbations are then computed
                    conditioned on this Bohmian trajectory. In practice, during the
                    contraction phase, the significant part of the power spectrum is
                    generated in the classical regime, effectively following a
                    semi-classical approach. Quantum effects on the background become
                    relevant only near the bounce, where the relevant modes are on
                    super-Hubble scales. If the background is in a superposition of
                    Gaussian states, the perturbations would naturally reflect this
                    superposition. However, this aspect lies outside the scope of our
                    work, as we are focused here on black hole formation during the
                    contraction phase.

@@

          \end{enumerate}

    \item In general, when both the background and the perturbations are quantized it is necessary
          to develop the second order Lagrangian for the perturbations without assuming the background
          is classical. Once this is done, one can show that for particular choices of the wave function
          for the whole system (background plus perturbations), the perturbations can be treated as
          quantum fields evolving around the Bohmian trajectory. We included a brief discussion of this
          point in the revised manuscript, Sec.~3.3.

@@

    \item In Ref.~\cite{vitenti2012large}, it is shown that for a Friedmann background with
          scalar perturbations, an appropriate gauge choice ensures the perturbative
          series remains valid in the bounce model. Nevertheless, this only shows that
          vacuum initial conditions around an initial Friedmann background are
          consistent with the perturbative expansion. The referee is correct in pointing
          out that the perturbative expansion may not be valid for more general initial
          conditions. In our study, we focus on this simpler model, leaving the
          exploration of a homogeneous and anisotropic background, specifically
          Bianchi-I models, for future work. A brief discussion of this possibility has
          been included in the conclusion.

    \item To simplify the presentation, we set the speed of light $c=1$ in the main
          text. We have now included a brief discussion of this choice in the revised
          manuscript. We also clarified the definition of energy density and perturbation
          in the main text.
\end{enumerate}

### 2024-11-19 main.tex
Checked grammar and c constant

	After performing a Legendre transformation and solving the constraints, the Lagrangian
	simplifies to:
	\begin{align}
		L          & = \Pi_a \dot{\bar{a}} + \Pi_{\varphi_1} \dot{\varphi}_1 +
		N \left( \frac{\kappa \Pi_a^2}{12 V \bar{a}} - V \bar{a}^3 \bar{\rho} \right), \\
		\Pi_a      & = -\frac{6 V \bar{a} \dot{\bar{a}}}{N \kappa}, \qquad
		\Pi_{\varphi_1} = V \bar{a}^3 f \frac{1+w}{w}\bar{v}^{1/w},                    \\
		\bar{\rho} & = \frac{f}{w\bar{a}^{3(1+w)}} \left( \frac{w}{1+w}
		\frac{1}{Vf} \right)^{1+w}\Pi_{\varphi_1}^{1+w},
	\end{align}
	where $V$ is the conformal volume. Since the fluid term has the form
	$\Pi_{\varphi_1}^{1+w}$ and the momentum term for the scale factor includes a factor
	dependent on $\bar{a}$, we introduce a final canonical transformation to define new
	momenta, $\Pi_q$ and $\Pi_T$.
	\begin{align}
		\Pi_q & = \Pi_{\bar{a}} \bar{a}^{(-1+3w)/2},                                                 \\
		\Pi_T & = \frac{V f}{w} \left( \frac{w}{1+w} \frac{1}{Vf}\right)^{1+w} \Pi_{\varphi_1}^{1+w}
	\end{align}

	In the above framework, the total Hamiltonian takes the form
	\begin{align}
		\label{total_hamiltonian_FLRW}
		{\cal H} & = \frac{N}{\bar{a}^{3w}}\dpar{\Pi_{T}- \frac{\kappa\Pi_q^{2} }{12}} ,
	\end{align}
	Canonical quantization is then performed by promoting classical variables to operators
	satisfying the canonical commutation relations. This process yields the following
	Wheeler-DeWitt equation for the wave-function of the universe  $\Psi(\bar{a}, T)$
	(see Ref.~\cite{nelson_peter_bouncing_original}):
	\begin{equation}
		\label{wdweq}
		i\hbar\frac{\partial}{\partial T}\Psi(q,T) -
		\frac{\kappa\hbar^2}{12}\frac{\partial^{2}}{\partial q^{2}}\Psi(q,T)  = 0 ,
	\end{equation}
	such that a specific operator factor ordering was chosen to preserve the symmetries of
	the classical system \cite{halliwell1990introductory}. Also, note that our last
	canonical transformation lead to the following variable
	\begin{equation}
		q = \frac{2\bar{a}^{\frac{3}{2}\dpar{1-w} } }{ 3\dpar{1-w } }.
	\end{equation}
	In the WdW equation above, we circumvent the problem of time by interpreting the
	parameter $T$ as an intrinsic time variable that is monotonically related to the
	classical cosmic time in the classical limit. This allows us to interpret the equation
	as a Schrödinger-like equation.
}

Equation~\eqref{wdweq} resembles a time-reversed free particle Schrödinger equation and,
with appropriate boundary conditions, its solutions are wave-functions of the scale
factor $\bar{a}$. We turn to the De Broglie-Bohm
interpretation~\cite{mukhanov2005physical, nelson2000bohm}, such that assuming a Gaussian
wave-function $\Psi(q, T)$, the Bohmian trajectory solution translates to the scale
factor reads
\begin{equation}
	\label{bohm_scale_factor}
	\bar{a}(T) = \bar{a}_{B}\dcol{1 + \dpar{ \frac{T}{T_{ B}} }^2 }^{{\frac{1}{3}\frac{1}{\dpar{1-w}} } }  ,
\end{equation}
where $\bar{a}_{B}$ is an integration constant that represents the minimum scale factor
value and $T_B$ is a small arbitrary constant related to the time scale of the bounce.
	{\color{red} To simplify the analysis, we choose the lapse function $N = \bar{a}^{3w}$,
		but this choice is only possible after solving the Bohmian trajectory. The reason is
		that, initially, we cannot select $N$ as an operator function of the scale factor,
		However, once we have the Bohmian trajectory $\bar{a}(T)$, we can make this choice,
		using it purely for convenience in the analysis. In our model, $T$ extends from $T =
			-\infty$ in the far past to $T = 0$ at the bounce, where the scale factor $\bar{a}$
		reaches its minimum. } Remarkably, $\bar{a}(T) \neq 0$ for all $T$, which means that
this model is non-singular and represents an eternal
universe~\cite{nelson_peter_bouncing_original}. For a detailed derivation, see
\cite{nelson2021bouncing}.

@@

In the above equations, there is an additional term
$-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation
for a classical universe. {\color{red} In our model, the scale factor dynamics resemble
		a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho} +
			\rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
			\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
		state $w_q = 1$ that accounts for quantum effects to avoid the primordial
		singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
		physical energy density but rather an effective one that emerges from the Bohmian. In
		the perturbative analysis, the energy density will always be associated only with
		$\bar{\rho}$. For large-scale factors $\bar{a} \gg \bar{a}_{b} $,
		Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the
		classical dust dominated flat FLRW universe. With this, we conclude the analysis of our
		background quantum bouncing model and move to its associated perturbations. In the next
		sections, we will consider the above quantum bouncing model with $w\approx 10^{-10}$ and
		$\bar{a}(t_0)/a_b = 10^{35}$ where $t_0$ is the time during the contracting phase where
		$H(t_0) = -H_0$ for $H_0 = 70 \, \text{km/s/Mpc}$. Moreover, when solving the equations
		numerically, we use the exact expressions in terms of $w$ without approximation.}

@@

We are mostly interested in the energy density perturbations that collapse to form PBHs.
These perturbations can be examined using the density contrast, defined as
\begin{align}
	\label{densitycon}
	\delta & \equiv \frac{\delta \rho}{\bar{\rho} + \bar{p}}
	,\end{align}
where $\delta\rho$ is the perturbation to the background energy density $\bar\rho$. The
density contrast provides a normalized measurement of the energy density perturbation
around the background matter density field. However, once again we are interested in its
gauge invariant form
\begin{align}
	\label{deltarhoinvariant}
	\tilde{\delta\rho} & \equiv \delta\rho - \mathcal{V} \dot{\Bar{\rho}}
	,\end{align}
where $\mathcal{V}$ is the velocity perturbation of the fluid, which has its gauge
invariant form
\begin{align}
	\tilde{\mathcal{V}} & \equiv \mathcal{V} + \delta\sigma
	.\end{align}

Using the background Einstein equations, we can relate the density contrast to the
Bardeen variables and the gauge invariant curvature perturbation $\zeta$ through the
following relations~\cite{Mukhanov1992, Vitenti2013}
\begin{align}
	\Psi                                                    & =\Phi,
	\\
	\label{deltarho}
	-\frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})} & =   \frac{{\tilde{\delta\rho}}} {3(\bar{\rho}+\bar{p})}
	,
	\\
	\label{vrelation}
	\zeta                                                   & \equiv \Psi  + \bar{H} \tilde{\mathcal{V}}
	\\
	\label{zeta2}
	\zeta                                                   & = \frac{3  \bar{a}^3}{N^2z^2 c_s^2 } \left[\frac{\partial}{\partial t}\left(\frac{c\Phi }{3\bar{H}}\right) + \frac{\Phi}{3}\right]
	,                                                                                                                                                                                            \\
	\label{phiz}
	\bar{D}^{2} \Phi                                        & =  - \frac{z^2\bar{H}}{\bar{a}^3} \dot{\zeta} = -\frac{\bar{H}}{2\bar{a}^3} \Pi_\zeta
	.\end{align}
\textcolor{red}{In the above, $c^{2}_s=\dpar{ \frac{\partial\bar{\rho}}{\partial\bar{p}} }_{S}= w$ is the speed of sound, $\Pi_\zeta$ is the conjugated momenta related to $\zeta$ and
	\begin{align}
		\label{zdef}
		z^2=\frac{\kappa \bar{a}^3 (\bar{\rho} + \bar{p})}{2\bar{H}^2 c_{s}^2}
		.\end{align}}
For completeness, we also write the relation between the curvature perturbation and the
usual Mukhanov-Sasaki variable $v$, that is,
\begin{align}
	\label{msv}
	v & \equiv - \zeta z \sqrt{\frac{2}{\kappa}}
	.\end{align}
We will establish a connection between PBH formation and the density contrast in the next section. For now, it suffices to understand that the excess energy density associated with the perturbations leads to black hole formation, and we can measure such excess through the density contrast. Thus our definition in Eq.~\eqref{deltarhoinvariant} is extremely important and we will now analyze it.

Note from Eq.~\eqref{deltarhoinvariant} that if we choose a gauge where $\mathcal{V} =0$, $\Tilde{\delta\rho}$ becomes the physical density perturbation, which leads to an easier interpretation of this quantity. Also, the choice of $\mathcal{V} =0$ will be well suited to connect our perturbation theory with the Lemaitre-Toman-Bondi metric discussed in App.~\ref{appgauge}.  Other works have defined the gauge invariant density perturbation as
\begin{align}
	\label{newtoniandeltarho}
	\Tilde{\delta\rho}^{N} & \equiv \delta\rho - \delta\sigma \dot{\Bar{\rho}}
	,\end{align}
so that this quantity becomes the physical variable in the Newtonian Gauge, where $\delta\sigma = 0$ and $\Phi = \phi$. However, in Ref.~\cite{vitenti2012large}, it was shown that for bounce models with long contracting phases, regardless of the bounce type, the Bardeen perturbation $\Phi$ grows larger than one and invalidates the perturbative series, as $\phi$ also grows larger than one in the Newtonian Gauge. Additionally, the definition in Eq.~\eqref{newtoniandeltarho} leads to
\begin{align}
	\zeta & = \Phi + \frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})}+  \frac{{\tilde{\delta\rho}^N}} {3(\bar{\rho}+\bar{p})}
	,\end{align}
which implies that $\tilde{\delta\rho}^N$ grows with $\Phi$ and thus has larger spectra
as well in this gauge. Hence, the Newtonian gauge is not a valid choice for bounce
models as it would lead to miss-calculation of the physical quantities whose values
would be inflated. We shall avoid this choice in this work and stick with the definition
in Eq.~\eqref{deltarhoinvariant}. We now compute the density contrast modes, which
require the perturbative equations of motion of the model.

@@

The perturbations are described by the Einstein-Hilbert action expanded up to second order, resulting in the Mukhanov-Sasaki Lagrangian
\begin{align}
	\label{lagms}
	L_{MS} & = \int \dd^3\textbf{x} \frac{1}{ \kappa} \left({\dot{\zeta}}^2 z^2 + c_s^2 z^2 \zeta\Delta \zeta\right)
	,\end{align}
where $\Delta = \bar{D}^2$ is the spatial Laplacian operator and $z$ is given by Eq.~\eqref{zdef}.
To compute the density contrast in Eq.~\eqref{deltarho}, we will need the curvature modes related to the Mukhanov-Sasaki Lagrangian. The extremization of Eq.~\eqref{lagms} in Fourier space yields the equation of motion
\begin{align}
	\label{four}
	\Ddot{\zeta}_{k}+ 2\frac{ \dot{z} }{ z } \dot{\zeta}_k + \frac{k^2}{\bar{a}^2} \zeta_k & =0
	,\end{align}
where $k$ is the comoving wave number of the modes. Since obtaining analytical solutions to \eqref{four} is non-trivial, we will employ a numerical code to compute the modes, which requires the use of dimensionless variables. In dimensionless units, denoted with a subscript $A$, we redefine our variables as\footnote{A derivative of a dimensionless variable will also be dimensionless. However, we use the same dot notation.}
\begin{align}
	\zeta_{k_A, A}   & \equiv \frac{\zeta_{k}}{\sqrt{ \kappa \hbar R_H}},           \\
	\Pi_{\zeta_k, A} & \equiv \frac{\Pi_{\zeta_k} \sqrt{R_H}}{\sqrt{\kappa \hbar}}, \\
	\label{kadm}
	k_A              & \equiv k R_H,                                                \\
	t_A              & \equiv \frac{t}{R_H}
	,\end{align}
where $l_p$ is the Planck length and $R_H = \frac{c}{\bar{H}_0}$. The equation of motion becomes
\begin{align}
	\label{eomosc}
	\dot{\Pi}_{k_A, A}+  \frac{2 c^2_s z^2 k_A^2}{\bar{a}^2  } \zeta_{k_A, A} & =0 \nonumber~~ or \\
	\dot{\Pi}_{k_A, A}+  2 \omega_{k_a}^2 z^2  \zeta_{k_A, A}                 & =0
	.\end{align}
Here, $\omega_A^2 \equiv \frac{c_s^2 k_a^2}{\bar{a}^2}$ is the dimensionless frequency and the dimensionless conjugated momenta are given by
\begin{align}
	\Pi_{k_A, A} & = 2 z^2 \dot{\zeta}_{k_A, A}
	.\end{align}
Eq.~\eqref{eomosc} can be interpreted as a harmonic oscillator with a time-dependent frequency $\omega_{k_a}(t_a)$ and mass $m(t_a) = 2z^2$. From here on, we are going to use the dimensionless variables with the same notation as described above.

@@

In this model, the background is characterized by a quantum contracting phase where an
effective quantum fluid dominates near the bounce. Consequently, we seek quantized
perturbations for consistency, although both theories can be viewed independently. In
particular, the use of quantized perturbations evolving on classical backgrounds has
been widely used since the results of Mukhanov and Chibisov \cite{mukhanov1981quantum}
and Hawking \cite{hawking1982quantum_fluctuations} to derive the power spectrum of
primordial perturbations, which are in turn used to describe the formation of structure
on our universe. {\color{red} In our model, we quantize both the background and the
		perturbations. Since we adopt the De Broglie-Bohm interpretation for the background,
		quantum effects are described in terms of the Bohmian scale factor, as shown in
		Eq.~\eqref{bohm_scale_factor}. This allows us to treat quantum perturbations evolving on
		this scale factor. To demonstrate the consistency of this approach,
		Ref.~\cite{Peter2005} derived the second-order Lagrangian for tensor perturbations
		without assuming classical equations of motion for the background. Furthermore, by using
		a factorized wave function for the perturbations and the background, the authors showed
		that these perturbations evolve conditioned on the Bohmian scale factor. In
		\cite{Vitenti2013}, the equivalent Lagrangian for scalar fluid perturbations was
		derived and is the one we use in this work.}

@@

Using these initial conditions for the modes, a well-defined vacuum state is obtained,
which is in turn used to construct the Fock Space by successive applications of the
creation operator $\hat{a}^{\dagger}_{k}$ and their linear combinations
\cite{mukhanov2007introduction}. Now that we have defined the appropriate initial
conditions for quantization, we may use them to solve the equations of motion in
Eq.~\eqref{eomosc} and obtain the mode functions, which we shall do numerically.
	{\color{red} Finally, it is important to note that the adiabatic vacuum prescription is
		consistent with the perturbative analysis of the model, as it is leads to small
		perturbations as discussed in Ref.~\cite{vitenti2012large}.}

>>>>

	After performing a Legendre transformation and solving the constraints, the Lagrangian
	simplifies to:
	\begin{align}
		L          & = \Pi_a \dot{\bar{a}} + \Pi_{\varphi_1} \dot{\varphi}_1 +
		N \left( \frac{\kappa \Pi_a^2}{12 V \bar{a}} - V \bar{a}^3 \bar{\rho} \right), \\
		\Pi_a      & = -\frac{6 V \bar{a} \dot{\bar{a}}}{N \kappa}, \qquad
		\Pi_{\varphi_1} = V \bar{a}^3 f \frac{1+w}{w}\bar{v}^{1/w},                    \\
		\bar{\rho} & = \frac{f}{w\bar{a}^{3(1+w)}} \left( \frac{w}{1+w}
		\frac{1}{Vf} \right)^{1+w}\Pi_{\varphi_1}^{1+w},
	\end{align}
	where $V$ is the conformal volume. Since the fluid term has the form
	$\Pi_{\varphi_1}^{1+w}$ and the momentum term for the scale factor includes a factor
	dependent on $\bar{a}$, we introduce a final canonical transformation to define new
	momenta, $\Pi_q$ and $\Pi_T$
	\begin{align}
		\Pi_q & = \Pi_{\bar{a}} \bar{a}^{(-1+3w)/2},                                                 \\
		\Pi_T & = \frac{V f}{w} \left( \frac{w}{1+w} \frac{1}{Vf}\right)^{1+w} \Pi_{\varphi_1}^{1+w}
	\end{align}.

	In the above framework, the total Hamiltonian takes the form
	\begin{align}
		\label{total_hamiltonian_FLRW}
		{\cal H} & = \frac{N}{\bar{a}^{3w}}\dpar{\Pi_{T}- \frac{\kappa\Pi_q^{2} }{12}} ,
	\end{align}
	Canonical quantization is then performed by promoting classical variables to operators
	satisfying the canonical commutation relations. This process yields the following
	Wheeler-DeWitt equation for the wave-function of the universe  $\Psi(\bar{a}, T)$
	(see Ref.~\cite{nelson_peter_bouncing_original}):
	\begin{equation}
		\label{wdweq}
		i\hbar\frac{\partial}{\partial T}\Psi(q,T) -
		\frac{\kappa\hbar^2}{12}\frac{\partial^{2}}{\partial q^{2}}\Psi(q,T)  = 0 ,
	\end{equation}
	such that a specific operator factor ordering was chosen to preserve the symmetries of
	the classical system \cite{halliwell1990introductory}. Also, note that our last
	canonical transformation leads to the following variable
	\begin{equation}
		q = \frac{2\bar{a}^{\frac{3}{2}\dpar{1-w} } }{ 3\dpar{1-w } }.
	\end{equation}
	In the WdW equation above, we circumvent the problem of time by interpreting the
	parameter $T$ as an intrinsic time variable that is monotonically related to the
	classical cosmic time in the classical limit. This allows us to interpret the equation
	as a Schrödinger-like equation.
}

Equation~\eqref{wdweq} resembles a time-reversed free particle Schrödinger equation and,
with appropriate boundary conditions, its solutions are wave functions of the scale
factor $\bar{a}$. We turn to the De Broglie-Bohm
interpretation~\cite{mukhanov2005physical, nelson2000bohm}, such that assuming a Gaussian
wave-function $\Psi(q, T)$, the Bohmian trajectory solution translates to the scale
factor reads
\begin{equation}
	\label{bohm_scale_factor}
	\bar{a}(T) = \bar{a}_{B}\dcol{1 + \dpar{ \frac{T}{T_{ B}} }^2 }^{{\frac{1}{3}\frac{1}{\dpar{1-w}} } }  ,
\end{equation}
where $\bar{a}_{B}$ is an integration constant that represents the minimum scale factor
value and $T_B$ is a small arbitrary constant related to the time scale of the bounce.
	{\color{red} To simplify the analysis, we choose the lapse function $N = \bar{a}^{3w}$,
		but this choice is only possible after solving the Bohmian trajectory. The reason is
		that, initially, we cannot select $N$ as an operator function of the scale factor,
		However, once we have the Bohmian trajectory $\bar{a}(T)$, we can make this choice,
		using it purely for convenience in the analysis. In our model, $T$ extends from $T =
			-\infty$ in the far past to $T = 0$ at the bounce, where the scale factor $\bar{a}$
		reaches its minimum. } Remarkably, $\bar{a}(T) \neq 0$ for all $T$, which means that
this model is non-singular and represents an eternal
universe~\cite{nelson_peter_bouncing_original}. For a detailed derivation, see
\cite{nelson2021bouncing}.

@@

In the above equations, {\color{red} the overdot represents the time derivative and} there is an additional term
$-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation
for a classical universe. {\color{red} In our model, the scale factor dynamics resemble
		a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho} +
			\rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
			\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
		state $w_q = 1$ that accounts for quantum effects to avoid the primordial
		singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
		physical energy density but rather an effective one that emerges from the Bohmian trajectory. In
		the perturbative analysis, the energy density will always be associated only with
		$\bar{\rho}$. For large-scale factors $\bar{a} \gg \bar{a}_{b} $,
		Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the
		classical dust dominated a flat FLRW universe. With this, we conclude the analysis of our
		background quantum bouncing model and move to its associated perturbations. In the next
		sections, we will consider the above quantum bouncing model with $w\approx 10^{-10}$ and
		$\bar{a}(t_0)/a_b = 10^{35}$ where $t_0$ is the time during the contracting phase where
		$H(t_0) = -H_0$ for $H_0 = 70 \, \text{km/s/Mpc}$. Moreover, when solving the equations
		numerically, we use the exact expressions in terms of $w$ without approximation.}

@@

We are mostly interested in the energy density perturbations that collapse to form PBHs.
These perturbations can be examined using the density contrast, defined as
\begin{align}
	\label{densitycon}
	\delta & \equiv \frac{\delta \rho}{\bar{\rho} + \bar{p}}
	,\end{align}
where $\delta\rho$ is the perturbation to the background energy density $\bar\rho$. The
density contrast provides a normalized measurement of the energy density perturbation
around the background matter density field. However, once again we are interested in its
gauge-invariant form
\begin{align}
	\label{deltarhoinvariant}
	\tilde{\delta\rho} & \equiv \delta\rho - \mathcal{V} \dot{\Bar{\rho}}
	,\end{align}
where $\mathcal{V}$ is the velocity perturbation of the fluid, which has its gauge
invariant form
\begin{align}
	\tilde{\mathcal{V}} & \equiv \mathcal{V} + \delta\sigma
	.\end{align}

Using the background Einstein equations, we can relate the density contrast to the
Bardeen variables and the gauge invariant curvature perturbation $\zeta$ through the
following relations~\cite{Mukhanov1992, Vitenti2013}
\begin{align}
	\Psi                                                    & =\Phi,
	\\
	\label{deltarho}
	-\frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})} & =   \frac{{\tilde{\delta\rho}}} {3(\bar{\rho}+\bar{p})}
	,
	\\
	\label{vrelation}
	\zeta                                                   & \equiv \Psi  + \bar{H} \tilde{\mathcal{V}}
	\\
	\label{zeta2}
	\zeta                                                   & = \frac{3  \bar{a}^3}{N^2z^2 c_s^2 } \left[\frac{\partial}{\partial t}\left(\frac{\Phi }{3\bar{H}}\right) + \frac{\Phi}{3}\right]
	,                                                                                                                                                                                            \\
	\label{phiz}
	\bar{D}^{2} \Phi                                        & =  - \frac{z^2\bar{H}}{\bar{a}^3} \dot{\zeta} = -\frac{\bar{H}}{2\bar{a}^3} \Pi_\zeta
	.\end{align}
\textcolor{red}{In the above, $c^{2}_s=\dpar{ \frac{\partial\bar{\rho}}{\partial\bar{p}} }_{S}= w$ is the speed of sound, $\Pi_\zeta$ is the conjugated momenta related to $\zeta$ and
	\begin{align}
		\label{zdef}
		z^2=\frac{\kappa \bar{a}^3 (\bar{\rho} + \bar{p})}{2\bar{H}^2 c_{s}^2}
		.\end{align}}
For completeness, we also write the relation between the curvature perturbation and the
usual Mukhanov-Sasaki variable $v$, that is,
\begin{align}
	\label{msv}
	v & \equiv - \zeta z \sqrt{\frac{2}{\kappa}}
	.\end{align}
In the next section, we will establish a connection between PBH formation and density contrast. For now, it suffices to understand that the excess energy density associated with the perturbations leads to black hole formation, and we can measure such excess through the density contrast. Thus our definition in Eq.~\eqref{deltarhoinvariant} is extremely important and we will now analyze it.

Note from Eq.~\eqref{deltarhoinvariant} that if we choose a gauge where $\mathcal{V} =0$, $\Tilde{\delta\rho}$ becomes the physical density perturbation, which leads to an easier interpretation of this quantity. Also, the choice of $\mathcal{V} =0$ will be well suited to connect our perturbation theory with the Lemaitre-Toman-Bondi metric discussed in App.~\ref{appgauge}.  Other works have defined the gauge invariant density perturbation as
\begin{align}
	\label{newtoniandeltarho}
	\Tilde{\delta\rho}^{N} & \equiv \delta\rho - \delta\sigma \dot{\Bar{\rho}}
	,\end{align}
so that this quantity becomes the physical variable in the Newtonian Gauge, where $\delta\sigma = 0$ and $\Phi = \phi$. However, in Ref.~\cite{vitenti2012large}, it was shown that for bounce models with long contracting phases, regardless of the bounce type, the Bardeen perturbation $\Phi$ grows larger than one and invalidates the perturbative series, as $\phi$ also grows larger than one in the Newtonian Gauge. Additionally, the definition in Eq.~\eqref{newtoniandeltarho} leads to
\begin{align}
	\zeta & = \Phi + \frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})}+  \frac{{\tilde{\delta\rho}^N}} {3(\bar{\rho}+\bar{p})}
	,\end{align}
which implies that $\tilde{\delta\rho}^N$ grows with $\Phi$ and thus has larger spectra
as well as in this gauge. Hence, the Newtonian gauge is not a valid choice for bounce
models as it would lead to miss-calculation of the physical quantities whose values
would be inflated. We shall avoid this choice in this work and stick with the definition
in Eq.~\eqref{deltarhoinvariant}. We now compute the density contrast modes, which
require the perturbative equations of motion of the model.

@@

The perturbations are described by the Einstein-Hilbert action expanded up to second order, resulting in the Mukhanov-Sasaki Lagrangian
\begin{align}
	\label{lagms}
	L_{MS} & = \int \dd^3\textbf{x} \frac{1}{ \kappa} \left({\dot{\zeta}}^2 z^2 + c_s^2 z^2 \zeta\Delta \zeta\right)
	,\end{align}
where $\Delta = \bar{D}^2$ is the spatial Laplacian operator and $z$ is given by Eq.~\eqref{zdef}.
To compute the density contrast in Eq.~\eqref{deltarho}, we will need the curvature modes related to the Mukhanov-Sasaki Lagrangian. The extremization of Eq.~\eqref{lagms} in Fourier space yields the equation of motion
\begin{align}
	\label{four}
	\Ddot{\zeta}_{k}+ 2\frac{ \dot{z} }{ z } \dot{\zeta}_k + \frac{k^2}{\bar{a}^2} \zeta_k & =0
	,\end{align}
where $k$ is the comoving wave number of the modes. Since obtaining analytical solutions to \eqref{four} is non-trivial, we will employ a numerical code to compute the modes, which requires the use of dimensionless variables. In dimensionless units, denoted with a subscript $A$, we redefine our variables as\footnote{A derivative of a dimensionless variable will also be dimensionless. However, we use the same dot notation.}
\begin{align}
	\zeta_{k_A, A}   & \equiv \frac{\zeta_{k}}{\sqrt{ \kappa \hbar R_H}},           \\
	\Pi_{\zeta_k, A} & \equiv \frac{\Pi_{\zeta_k} \sqrt{R_H}}{\sqrt{\kappa \hbar}}, \\
	\label{kadm}
	k_A              & \equiv k R_H,                                                \\
	t_A              & \equiv \frac{t}{R_H}
	,\end{align}
where $l_p$ is the Planck length and $R_H = \frac{1}{\bar{H}_0}$. The equation of motion becomes
\begin{align}
	\label{eomosc}
	\dot{\Pi}_{k_A, A}+  \frac{2 c^2_s z^2 k_A^2}{\bar{a}^2  } \zeta_{k_A, A} & =0 \nonumber~~ or \\
	\dot{\Pi}_{k_A, A}+  2 \omega_{k_a}^2 z^2  \zeta_{k_A, A}                 & =0
	.\end{align}
Here, $\omega_A^2 \equiv \frac{c_s^2 k_a^2}{\bar{a}^2}$ is the dimensionless frequency and the dimensionless conjugated momenta are given by
\begin{align}
	\Pi_{k_A, A} & = 2 z^2 \dot{\zeta}_{k_A, A}
	.\end{align}
Eq.~\eqref{eomosc} can be interpreted as a harmonic oscillator with a time-dependent frequency $\omega_{k_a}(t_a)$ and mass $m(t_a) = 2z^2$. From here on, we are going to use the dimensionless variables with the same notation as described above.

@@

In this model, the background is characterized by a quantum contracting phase where an
effective quantum fluid dominates near the bounce. Consequently, we seek quantized
perturbations for consistency, although both theories can be viewed independently. In
particular, the use of quantized perturbations evolving on classical backgrounds has
been widely used since the results of Mukhanov and Chibisov \cite{mukhanov1981quantum}
and Hawking \cite{hawking1982quantum_fluctuations} to derive the power spectrum of
primordial perturbations, which are in turn used to describe the formation of structure
in our universe. {\color{red} In our model, we quantize both the background and the
		perturbations. Since we adopt the De Broglie-Bohm interpretation for the background,
		quantum effects are described in terms of the Bohmian scale factor, as shown in
		Eq.~\eqref{bohm_scale_factor}. This allows us to treat quantum perturbations evolving on
		this scale factor. To demonstrate the consistency of this approach,
		Ref.~\cite{Peter2005} derived the second-order Lagrangian for tensor perturbations
		without assuming classical equations of motion for the background. Furthermore, by using
		a factorized wave function for the perturbations and the background, the authors showed
		that these perturbations evolve conditioned on the Bohmian scale factor. In
		\cite{Vitenti2013}, the equivalent Lagrangian for scalar fluid perturbations was
		derived and is the one we use in this work.}

@@

Using these initial conditions for the modes, a well-defined vacuum state is obtained,
which is in turn used to construct the Fock Space by successive applications of the
creation operator $\hat{a}^{\dagger}_{k}$ and their linear combinations
\cite{mukhanov2007introduction}. Now that we have defined the appropriate initial
conditions for quantization, we may use them to solve the equations of motion in
Eq.~\eqref{eomosc} and obtain the mode functions, which we shall do numerically.
	{\color{red} Finally, it is important to note that the adiabatic vacuum prescription is
		consistent with the perturbative analysis of the model, as it leads to small
		perturbations as discussed in Ref.~\cite{vitenti2012large}.}
